---
title: "Mantel Test for correlation of two feature sets of single cell data sets"
author: "Yamada Ryo"
date: "2020/11/4"
output: html_document
---

## Input data sets

Let's assume we have two data sets, X and Y, that are the data record matrices of single cell measurementes.

Both are already prepared for UMAP, X and Y.

X and Y are number-of-sample x number-of-feature matrices and the order of single cells are matched between X and Y.

Let's assume $k_X, k_Y$ are already given so that they produce appropriate dimension reduction maps for X and Y, respectively.

Random data sets are generated below.

Appropriate $k_X,k_Y$ are somehow given.
```{r}
n.cell <- 100
n.feature_X <- 50
n.feature_Y <- 100
X <- matrix(rnorm(n.cell*n.feature_X),ncol=n.feature_X)
Y <- matrix(rnorm(n.cell*n.feature_Y),ncol=n.feature_Y)
k_X <- 10
k_Y <- 20
```


```{r}
library(umap)
library(vegan)
library(igraph)

# X and Y are well-prepared and given
# k_X, k_Y are given
# UMAP
umap_X <- umap(X,n_neighbors=k_X)
umap_Y <- umap(Y,n_neighbors=k_Y)

# Take out knn-graph object
knn_X <- umap_X$knn 
knn_Y <- umap_Y$knn

# Make edge lists
edge.listX <- edge.listY <- matrix(0,0,2)

for(i in 1:n.cell){
  edge.listX <- rbind(edge.listX,cbind(rep(i,k_X-1),knn_X[[1]][i,2:k_X]))
  edge.listY <- rbind(edge.listY,cbind(rep(i,k_Y-1),knn_Y[[1]][i,2:k_Y]))
}

# Make a graph object of igraph
gX <- graph.edgelist(edge.listX,directed=FALSE)
gY <- graph.edgelist(edge.listY,directed=FALSE)

# Graph distance matrix
dist.gX <- distances(gX,weights = c(t(knn_X[[2]][,2:k_X])))
dist.gY <- distances(gY,weights = c(t(knn_Y[[2]][,2:k_Y])))

# Mantel test
mantel.out <- mantel(dist.gX,dist.gY,method="spearman",permutations=1000,na.rm=TRUE)

print(mantel.out)
```
```{r}
# Correlation coefficient distribution
hist(mantel.out$perm)
```

