---
title: "Jordan Curve"
author: "ryamada"
date: "2020/5/29"
output: html_document
---

## Jordan Curve

Jordan curves are the simple closed curves , or closed curves without self-crossings, in 2D plane.

## Generation of Closed Curve

A closed curve is represented as a complex function as below.

Real part and imaginary part are both consisted of trigonometric functions.

In other words, any closed curves can be represented by two Fourrier series.

$$
z(t) = x(t) + i \times y(t), \; t \in [0,1)
$$


```{r}
freq.re <- 1:4
freq.im <- 1:4
a.re <- rnorm(length(freq.re))
a.im <- rnorm(length(freq.im)) 
ph.re <- rnorm(length(freq.re)) 
ph.im <- rnorm(length(freq.im)) 

t <- seq(from=0, to=1,length=201) * 2 * pi
t <- t[-1]

x <- y <- rep(0,length(t))

for(i in 1:length(freq.re)){
  x <- x + a.re[i] * sin(freq.re[i] * t + ph.re[i])
}
for(i in 1:length(freq.im)){
  y <- y + a.im[i] * sin(freq.im[i] * t + ph.im[i])
}

plot(x,y,type="l")
```

## Two points on the curve and their arc distance (geodesic distance) and euclidean distanfce in 2D plane

```{r}
diff.x <- diff(x[c(1:length(x),1)])
diff.y <- diff(y[c(1:length(y),1)])

len.segments <- sqrt(diff.x^2+diff.y^2)

```

```{r}
total.len <- sum(len.segments)
D.geod <- matrix(0,length(t),length(t))
for(i in 1:(length(t)-1)){
  for(j in (i+1):length(t)){
    tmp <- sum(len.segments[i:j])
    # There are two paths
    # i -> i+1 -> ... -> j-1 -> j : tmp
    # i -> i-1 -> ... -> j+1 -> j : total.len -tmp
    # shorter length of these two is the geodesic dist
    D.geod[i,j] <- D.geod[j,i] <- min(tmp,total.len-tmp)
  }
}
```

```{r}
D.Euc <- matrix(0,length(t),length(t))
for(i in 1:(length(t)-1)){
  for(j in (i+1):length(t)){
    tmp <- sum(len.segments[i:j])
    D.Euc[i,j] <- D.Euc[j,i] <- sqrt((x[i]-x[j])^2+(y[i]-y[j])^2)
  }
}
```
```{r}
par(mfcol=c(1,2)) 
image(D.Euc)
image(D.geod)
par(mfcol=c(1,1))
```

```{r}
plot(c(D.Euc),c(D.geod),pch=20,cex=0.1)
```