---
title: "LU decomposition and calculation complexity"
author: "ryamada"
date: "2020/5/26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## $1^2+2^2 +... = \sum_{i=1}^n i^2$

Let's calculate for n = 1,2,3...

```{r}
ns <- 1:100

vs <- rep(0,length(ns))

for(i in 1:length(ns)){
  vs[i] <- sum((1:i)^2)
}

plot(ns,vs)
plot(ns,log10(vs),ylab="log10")
```

## How about $\frac{1}{3} n^3

```{r}
ws <- rep(0,length(ns))
for(i in 1:length(ns)){
  ws[i] <- 1/3 * i^3
}
plot(ns,ws)
plot(ns,log10(ws),ylab="log10")
```

## Plot two value series together

They are essentially the same.

```{r}
vws <- cbind(vs,ws)

matplot(ns,vws,type="l")
matplot(ns,log10(vws),type="l",ylab="log10")
```

## Addendum

The sequence of integers can be checked @ https://oeis.org/?language=chineseS .

$\sum_{i=1}^n i^2$ is;
```{r}
vs
```

Input the initial several values 1,5,14,...,into the query box of the site.

It returns the sequence ID https://oeis.org/A000330 .

It tells you that $\sum_{i=0}^n i^2 = 0^2 + 1^2 + ... + = n(n+1)(2n+1)/6$

You can double check this formula as below.

```{r}
vs2 <- rep(0,length(ns)) 
for(i in 1:length(ns)){
  vs2[i] <- i*(i+1)*(2*i+1)/6
}
vs2
plot(vs,vs2)
abline(0,1,col=2)
```

How about $\frac{1}{3}n^3$? The sequence is not integers, so the integer sequence site can not be used.

## Compare the formulae

Let's compare two formulae; $n(n+1)(2n+1)/6$ and $1/3 n^3$

$$
n(n+1)(2n+1)/6 = (2n^3 + 3n^2 + n)/6 = n^3/3 + n^2/2 + n/6
$$

Because $n^2/2 + n/6$ gets too small compared to $n^3/3$, when n gets big, therefore the number of steps is said to increase as like $n^3/3$.


